OVERVIEW RESULTS: Harmonized invertebrate grouping feature databases

### 2020-08-31 Paper/Results/SI.tex
31.08: Intro+Methods+Results update

\section*{Supporting Information}

@@

\subsection*{Re-analysis of Szöcs et al. 2014 using harmonized and aggregated grouping features.}
\label{subsec:SI_szoecs_reanalysis}

@@

\section*{Harmonization of the European trait databases}
\label{sec:SI_harmonization_EU}

% \cellcolor{blue!25}
\begin{longtable}{llll}
%\begin{table}[ht]
%    \centering
    \caption{Representation of traits per grouping feature for the used trait databases. The color coding indicates traits that have been harmonized. Cyan colored traits have not been used because they either represented ambiguous traits or traits that were not compatible with the traits of the other databases. Harmonization was done by assigning the maximum affinity of the allocated traits for the respective taxa to the harmonized trait.}
    \endfirsthead
    \endhead
    %\begin{tabular}{llll}
    \hline
    Grouping feature & Freshwater ecology & Tachet & Harmonized traits \\ 
    \hline
    \hline
    Voltinism & Semivoltine & Semivoltine & Semivoltine \\ 
    Voltinism & Univoltine & Univoltine & Univoltine \\ 
    \rowcolor{gray!25}
    Voltinism & Bivoltine & Polyvoltine & Bi/Multivoltine \\ 
    \cellcolor{gray!25}Voltinism & \cellcolor{gray!25}Trivoltine &  &  \\ 
    \cellcolor{gray!25}Voltinism & \cellcolor{gray!25}Multivoltine &  &  \\ 
    \cellcolor{gray!25}Voltinism & \cellcolor{gray!25}Flexible &  &  \\ 
    \rowcolor{blue!25}
    \cellcolor{blue!25}Feeding Mode & \cellcolor{blue!25}Shredder & Shredder & Shredder \\ 
    \cellcolor{blue!25}Feeding Mode & \cellcolor{blue!25}Miner & Deposit-feeder & Gatherer \\ 
    \cellcolor{blue!25}Feeding Mode & \cellcolor{blue!25}Xylophagus & \cellcolor{blue!50}Absorber & \cellcolor{blue!50}Filterer \\ 
    Feeding Mode & Gatherer & \cellcolor{blue!50}Filter-feeder & Herbivore \\ 
    \cellcolor{blue!50}Feeding Mode & \cellcolor{blue!50}Active filterer & Scraper & Predator \\ 
    \cellcolor{blue!50}Feeding Mode & \cellcolor{blue!50}Passive filterer & Predator & Parasite \\ 
    Feeding Mode & Grazer & Parasite &  \\ 
    Feeding Mode & Predator & \color{cyan}Piercer (plants or animals) &  \\ 
    Feeding Mode & Parasite &  &  \\ 
    \color{cyan}Feeding Mode & \color{cyan}Other &  &  \\ 
    \hline
    \rowcolor{green!25}
    Locomotion & Swimming/scating & Surface swimmer & Swimmer \\ 
    \cellcolor{green!25}Locomotion & \cellcolor{green!25}Swimming/diving & \cellcolor{green!25}Full water swimmer & Burrower \\ 
    Locomotion & Burrowing/boring & Burrower & Crawler \\ 
    Locomotion & Sprawling/walking & Crawler & \cellcolor{green!50}Sessil \\ 
    \cellcolor{green!50}Locomotion & \cellcolor{green!50}(semi) sessil & \cellcolor{green!50}Temporarily attachted &  \\ 
    \color{cyan}Locomotion & \color{cyan}Other & \cellcolor{green!50}Permanently attached &  \\ 
    \color{cyan}Locomotion &  & \color{cyan}Interstitial &  \\ 
    \color{cyan}Locomotion &  & \color{cyan}Flier &  \\
    \hline
    Respiration & Tegument & Tegument & Tegument \\ 
    Respiration & Gill & Gill & Gills \\ 
    \rowcolor{orange!25}
    Respiration & Plastron & Plastron & Plastron, spiracle \\ 
    \cellcolor{orange!25}Respiration & \cellcolor{orange!25}Spiracle (aerial) & \cellcolor{orange!25}Spiracle (aerial) &  \\ 
    \color{cyan}Respiration & \color{cyan}Hydrostatic vesicle & \color{cyan}Hydrostatic vesicle (aerial) &  \\ 
    \color{cyan}Respiration & \color{cyan}Tapping (air stores of aq. plants) &  &  \\ 
    \color{cyan}Respiration & \color{cyan}Excursion/Extension (to surface) &  &  \\ 
    \hline
    Body size &  & $<$= 0.25cm & Small ($<$ 1 cm) \\ 
    Body size &  & $>$ 0.25 - 0.5cm & Medium ($>$= 1cm - 2 cm) \\ 
    Body size &  & $>$ 0.5- 1cm & Large ($>$= 2 cm) \\ 
    Body size &  & $>$ 1 – 2 cm &  \\ 
    Body size &  & 2 – 4 cm &  \\ 
    Body size &  & 4 – 8 cm &  \\ 
    Body size &  & $>$ 8 cm &  \\ 
    \hline
    Reproduction & ovovivipar & ovoviviparity & ovoviviparity \\ 
    \rowcolor{purple!25}
    Reproduction & free isolated eggs & isolated eggs, free & aquatic eggs \\ 
    \cellcolor{purple!25}Reproduction & \cellcolor{purple!25}cemented isolated eggs & \cellcolor{purple!25}isolated eggs, cemented & terrestrial eggs \\
    \cellcolor{purple!25}Reproduction & \cellcolor{purple!25}fixed clutches & \cellcolor{purple!25}clutches, cemented or fixed &  \\ 
    \cellcolor{purple!25}Reproduction & \cellcolor{purple!25}free clutches & \cellcolor{purple!25}clutches, free &  \\ 
    \cellcolor{purple!25}Reproduction & \cellcolor{purple!25}clutches in vegetation & \cellcolor{purple!25}clutches, in vegetation &  \\ 
    Reproduction & terrestrial clutches & clutches, terrestrial &  \\ 
    \color{cyan}Reproduction & \color{cyan}asexual & \color{cyan}asexual reproduction &  \\ 
    \color{cyan}Reproduction & \color{cyan}parasitic &  &  \\ 
    \hline
    %\end{tabular}
%\end{table}
\end{longtable}

>>>>

\section*{Supporting Information}
\label{sec:SI}

@@

\subsection*{Re-analysis of Szöcs et al. 2014 using harmonized grouping features.}
\label{subsec:SI_szoecs_reanalysis}

@@

% \section*{Harmonization of the European trait databases}
% \label{sec:SI_harmonization_EU}

% % \cellcolor{blue!25}
% \begin{longtable}{lll|l}
% %\begin{table}[ht]
% %    \centering
%     \caption{Representation of traits per grouping feature and their harmonization for the European trait databases. The color coding indicates traits that have been harmonized. Cyan colored traits have not been used because they were too different from the amalgamated traits.}
%     \endfirsthead
%     \endhead
%     %\begin{tabular}{llll}
%     \hline
%     Grouping feature & Freshwater ecology & Tachet & Amalgamated traits \\ 
%     \hline
%     \hline
%     Voltinism & Semivoltine & Semivoltine & Semivoltine \\ 
%     Voltinism & Univoltine & Univoltine & Univoltine \\ 
%     \rowcolor{gray!25}
%     Voltinism & Bivoltine & Polyvoltine & Bi/Multivoltine \\ 
%     \cellcolor{gray!25}Voltinism & \cellcolor{gray!25}Trivoltine &  &  \\ 
%     \cellcolor{gray!25}Voltinism & \cellcolor{gray!25}Multivoltine &  &  \\ 
%     \cellcolor{gray!25}Voltinism & \cellcolor{gray!25}Flexible &  &  \\ 
%     \rowcolor{blue!25}
%     \cellcolor{blue!25}Feeding Mode & \cellcolor{blue!25}Shredder & Shredder & Shredder \\ 
%     \cellcolor{blue!25}Feeding Mode & \cellcolor{blue!25}Miner & Deposit-feeder & Gatherer \\ 
%     \cellcolor{blue!25}Feeding Mode & \cellcolor{blue!25}Xylophagus & \cellcolor{blue!50}Absorber & \cellcolor{blue!50}Filterer \\ 
%     Feeding Mode & Gatherer & \cellcolor{blue!50}Filter-feeder & Herbivore \\ 
%     \cellcolor{blue!50}Feeding Mode & \cellcolor{blue!50}Active filterer & Scraper & Predator \\ 
%     \cellcolor{blue!50}Feeding Mode & \cellcolor{blue!50}Passive filterer & Predator & Parasite \\ 
%     Feeding Mode & Grazer & Parasite &  \\ 
%     Feeding Mode & Predator & \textit{Piercer (plants or animals)} \textsuperscript{\textit{a}} &  \\ 
%     Feeding Mode & Parasite &  &  \\ 
%     \color{cyan}Feeding Mode & \color{cyan}Other &  &  \\ 
%     \hline
%     \rowcolor{green!15}
%     Locomotion & Swimming/scating & Surface swimmer & Swimmer \\ 
%     \cellcolor{green!15}Locomotion & \cellcolor{green!15}Swimming/diving & \cellcolor{green!15}Full water swimmer & \cellcolor{green!50}Burrower \\ 
%     \cellcolor{green!50}Locomotion & \cellcolor{green!50}Burrowing/boring & \cellcolor{green!50}Burrower & Crawler \\
%     Locomotion &  & \cellcolor{green!50}Interstitial & \cellcolor{green!95}Sessil \\
%     Locomotion & Sprawling/walking & Crawler &  \\ 
%     \cellcolor{green!95}Locomotion & \cellcolor{green!95}(semi) sessil & \cellcolor{green!95}Temporarily attachted &  \\ 
%     \color{cyan}Locomotion & \color{cyan}Other & \cellcolor{green!95}Permanently attached &  \\ 
%     \color{cyan}Locomotion &  & \color{cyan}Flier &  \\
%     \hline
%     Respiration & Tegument & Tegument & Tegument \\ 
%     Respiration & Gill & Gill & Gills \\ 
%     \rowcolor{orange!25}
%     Respiration & Plastron & Plastron & Plastron, spiracle \\ 
%     \cellcolor{orange!25}Respiration & \cellcolor{orange!25}Spiracle (aerial) & \cellcolor{orange!25}Spiracle (aerial) &  \\ 
%     \color{cyan}Respiration & \color{cyan}Hydrostatic vesicle & \color{cyan}Hydrostatic vesicle (aerial) &  \\ 
%     \color{cyan}Respiration & \color{cyan}Tapping (air stores of aq. plants) &  &  \\ 
%     \color{cyan}Respiration & \color{cyan}Excursion/Extension (to surface) &  &  \\ 
%     \hline
%     Body size &  & $<$= 0.25cm & Small ($<$ 1 cm) \\ 
%     Body size &  & $>$ 0.25 - 0.5cm & Medium ($>$= 1cm - 2 cm) \\ 
%     Body size &  & $>$ 0.5- 1cm & Large ($>$= 2 cm) \\ 
%     Body size &  & $>$ 1 – 2 cm &  \\ 
%     Body size &  & 2 – 4 cm &  \\ 
%     Body size &  & 4 – 8 cm &  \\ 
%     Body size &  & $>$ 8 cm &  \\ 
%     \hline
%     Reproduction & ovovivipar & ovoviviparity & ovoviviparity \\ 
%     \rowcolor{purple!25}
%     Reproduction & free isolated eggs & isolated eggs, free & aquatic eggs \\ 
%     \cellcolor{purple!25}Reproduction & \cellcolor{purple!25}cemented isolated eggs & \cellcolor{purple!25}isolated eggs, cemented & terrestrial eggs \\
%     \cellcolor{purple!25}Reproduction & \cellcolor{purple!25}fixed clutches & \cellcolor{purple!25}clutches, cemented or fixed &  \\ 
%     \cellcolor{purple!25}Reproduction & \cellcolor{purple!25}free clutches & \cellcolor{purple!25}clutches, free &  \\ 
%     \cellcolor{purple!25}Reproduction & \cellcolor{purple!25}clutches in vegetation & \cellcolor{purple!25}clutches, in vegetation &  \\ 
%     Reproduction & terrestrial clutches & clutches, terrestrial &  \\ 
%     \color{cyan}Reproduction & \color{cyan}asexual & \color{cyan}asexual reproduction &  \\ 
%     \color{cyan}Reproduction & \color{cyan}parasitic &  &  \\ 
%     \hline
%     %\end{tabular}
% %\end{table}
% \end{longtable}
% \begin{minipage}{\linewidth}\small
%     \textit{a} Taxa exhibiting this trait have been assigned to predators or herbivores based on a classification by Philippe Usseglio-Polatera.
% \end{minipage}

### 2020-08-31 Paper/Results/Taxonomic_cov_and_info_compl.tex
31.08: Intro+Methods+Results update

% \usepackage[utf8]{inputenc}
% \usepackage{hyperref}
% \usepackage{float}
% \usepackage[table,xcdraw]{xcolor}
% \usepackage{color, colortbl}
% \usepackage{longtable}
% %\usepackage[sort&compress,square,comma,authoryear]{natbib}
% \usepackage{booktabs}
% \usepackage{graphicx}
% \graphicspath{{/home/kunz/Dokumente/Projects/Trait_DB/Invertebrate_traits/Paper/Figures/}}
% \usepackage{longtable}
% \usepackage{rotating}
% \usepackage{geometry}
% \usepackage{array}

% \definecolor{Gray}{gray}{0.9}

% \newcommand{\specialcell}[2][c]{%
%   \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

@@

\section{Taxonomic coverage of the trait databases}

Regarding the taxonomical coverage, the New Zealand database has, as expected, the smallest taxon pool (Table \ref{tab:tax_coverage}). By contrast, the largest taxon pool is spanned by the European trait database with 4225 taxa followed by the North American trait database that contained trait information on 3542 taxa. The Australian database contains 1404 taxa. The European, New Zealand, and North American databases have most taxa on the highest taxonomical resolution while the Australian database has a similar number of taxa on species and genus-level.

% \begin{table}[ht]
%   \centering
%   \caption{Number of taxa per harmonized database and per taxonomic level. Numbers in parenthesis show relative frequencies in percentage.}
%   \label{tab:tax_coverage}
%   \begin{tabular}{lccccc}
%     \hline
%   Database & Nr. of taxa & Species & Genus & Family & Aquatic insects \\ 
%   \hline
%   EU & 4225 & 3951 (93.51) & 257 (6.08) & 17 (0.4) & 3654 (86.49) \\ 
%   NOA & 3542 & 2418 (68.27) & 1074 (30.32) & 50 (1.41) & 3144 (88.76) \\ 
%   AUS & 1402 & 564 (40.23) & 578 (41.23) & 260 (18.54) & 1015 (72.4) \\ 
%   NZ & 478 & 404 (84.52) & 47 (9.83) & 27 (5.65) & 443 (92.68) \\ 
%   \hline
%   \end{tabular}
%   \end{table}

\begin{table}[ht]
    \centering
    \caption{Number of taxa per harmonized database and per taxonomic level. Numbers in parenthesis show relative frequencies in percentage.}
    \label{tab:tax_coverage}
    \begin{tabular}{lccccc}
    \hline
    Database & Nr. of taxa & Species & Genus & Family & Nr. aquatic taxa \\ 
    \hline
    Europe & 4110 & 3848 (93.63) & 237 (5.77) & 25 (0.61) & 3579 (87.08) \\ 
      North America & 3753 & 2414 (64.32) & 1163 (30.99) & 176 (4.69) & 3305 (88.06) \\ 
      Australia & 1402 & 564 (40.23) & 578 (41.23) & 260 (18.54) & 1016 (72.47) \\ 
      New Zealand & 478 & 404 (84.52) & 47 (9.83) & 27 (5.65) & 443 (92.68) \\ 
    \hline
    \end{tabular}
\end{table}

\newpage

\section{Completeness of trait information}

The amount of entries with available information for the individual grouping features varied strongly for the European, North American and Australian databases (Table \ref{tab:trait_coverage}). By contrast, the New Zealand database contained complete trait information for most of the investigated grouping features (between 94 \% and 100 \%).

\begin{table}[ht]
    \centering
    \caption{Displayed is the percentage of entries that have information for the individual grouping features per database.} 
    \label{tab:trait_coverage}
    \begin{tabular}{llllllll}
    \hline
    Database & Body form & Oviposition & Voltinism & Locomotion & Size & Respiration & Feeding mode \\ 
    \hline
    EU & 7 & 16 & 24 & 33 & 11 & 56 & 65 \\ 
    NOA & 26 & 12 & 47 & 51 & 75 & 44 & 61 \\ 
    AUS & 5 & 48 & 51 & 42 & 78 & 70 & 99 \\ 
    NZ & 100 & 94 & 100 & 99 & 100 & 100 & 99 \\ 
    \hline
    \end{tabular}
    \end{table}

>>>>

@@

\subsection*{Taxonomic coverage of the harmonized trait datasets}

Regarding the taxonomical coverage, the New Zealand dataset has, as expected, the smallest taxon pool (478 taxa, Table \ref{tab:tax_coverage}). By contrast, the largest taxon pool is spanned by the European trait dataset with 4110 taxa followed by the North American trait dataset that contained trait information on 3753 taxa. The Australian dataset contains 1402 taxa. The European, New Zealand, and North American datasets have most taxa on the highest taxonomical resolution while the Australian dataset has a similar number of taxa on species and genus-level.

\begin{table}[ht]
    \centering
    \caption{Number of taxa per harmonized database and per taxonomic level. Numbers in parenthesis show relative frequencies in percentage.}
    \label{tab:tax_coverage}
    \begin{tabular}{lccccc}
    \\
    Database & Nr. of taxa & Species & Genus & Family & Nr. aquatic taxa \\ 
    \toprule[.1em]
    EU & 4110 & 3848 (93.63) & 237 (5.77) & 25 (0.61) & 3579 (87.08) \\ 
    NOA & 3753 & 2414 (64.32) & 1163 (30.99) & 176 (4.69) & 3305 (88.06) \\ 
    AUS & 1402 & 564 (40.23) & 578 (41.23) & 260 (18.54) & 1016 (72.47) \\ 
    NZ & 478 & 404 (84.52) & 47 (9.83) & 27 (5.65) & 443 (92.68) \\ 
    \bottomrule
    \end{tabular}
\end{table}

\subsection*{Completeness of trait information}

The amount of entries with available information for the selected grouping features varied strongly for the European, North American, and Australian datasets (Table \ref{tab:trait_coverage}). By contrast, the New Zealand dataset contained complete trait information for most of the investigated grouping features (between 94 \% and 100 \%).

\begin{table}[ht]
    \centering
    \caption{Displayed is the percentage of entries that have information for the individual grouping features per database.} 
    \label{tab:trait_coverage}
    \begin{tabular}{llllllll}
    \\
    Database & Body form & Oviposition & Voltinism & Locomotion & Size & Respiration & Feeding mode \\ 
    \toprule[.1em]
    EU & 7 & 16 & 24 & 33 & 11 & 56 & 65 \\ 
    NOA & 28 & 13 & 47 & 52 & 73 & 44 & 63 \\ 
    AUS & 5 & 48 & 51 & 42 & 78 & 70 & 99 \\ 
    NZ & 100 & 94 & 100 & 99 & 100 & 100 & 99 \\ 
    \bottomrule
    \end{tabular}
    \end{table}
